Mental Simulation as Bayesian Quadrature

### 2013-05-31 man/nips-2013.tex
Add citations, background, conclusion

% I think the main thing would be doing something more sophisticated
% in the search, rather than using the gradient. A less sophisticated
% adaptive sampling method seems like it would be best, and using von
% Mises makes sense.

@@

\TODO{Want to talk about previous work on strategies people might use,
  e.g. eye-tracking results from Just and Carpenter (1975?)}

% 1. The big picture
% 2. Our contribution
% 3. How we do it
% 4. What's in the paper

% The introduction could be made more general -- start with the
% question of how a rational agent should use her computational
% resources, and then get onto mental simulation/rotation as a special
% case of that, rather than starting with the phenomena. I would make
% the big picture goals clear at the start, and then segue into the
% mystery of mental rotation -- how do we know which way to rotate?

@@

We assume that mental imagery is indeed a visual process, and turn to
the question of how, and when, it is used. Previous work has examined
how people might use imagery \cite{Hegarty2004, Schwartz1999} to solve
reasoning problems, and when they might use imagery as opposed to a
symbolic rule \cite{Schwartz:1996uy}. People seem to use imagery
methodically, which raises the question: what is the method?

@@

We investigated strategies for performing the mental rotation task
\cite{Shepard1971} in two dimensions, and found that a nonparametric
model that enforces a positivity constraint on the likelihood function
performs best, as it is able to actively monitor the confidence of its
estimate. The simpler models performed much worse: they did not
maintain the linear relationship with the true angle of rotation due
to lack of robustness against local maxima, and they were inaccurate
at identifying whether the two shapes were the same or not.

One option for improving the performance of the simple models would be
to fit a threshold value, below which local maxima would be
ignored. However, this strategy would be rather brittle, for if the
distribution of shapes changed, the threshold would have to be
re-learned. It is possible that people exhibit this behavior, but we
cannot make any assumptions one way or the other without empirical
data.

A further strength of the nonparametric model over the other models,
however, is that it is likely to generalize well to three
dimensions. One aspect of \cite{Osborne:2012tm} which we have not yet
explored is their main contribution of \textit{active sampling}, in
which samples are iteratively selected to maximally decrease the
expected variance of $Z$. Due to the sequential constraint of mental
rotation, this strategy is not particularly useful in two dimensions,
as there are only every two directions in which to rotate. In three
dimensions, however, there are an infinite number of directions that
could be chosen after every step.

We conclude that, from this initial survey, models of mental rotation
which take an ``active'', directed approach seem well-suited to
explaining human behavior in these tasks. Future work will collect
empirical data from participants to perform a more quantitative
analysis of these modeling tools.

>>>>

@@

@@

We assume that mental imagery is indeed a visual process, and turn to
the question of how, and when, it is used. Previous work has examined
how people might use imagery \cite{Hegarty2004, Schwartz1999} to solve
reasoning problems, and when they might use imagery as opposed to a
symbolic rule \cite{Schwartz:1996uy}. Eye-tracking studies of people
performing mental rotations suggest strategies that people might be
using, such as aligning smaller features of the images
\cite{Just1976}. People seem to use imagery methodically, which raises
the question: what is the method?

@@

In this paper, we asked: how do people use mental simulation?  We
investigated the specific case of mental rotation, taking a rational
analysis approach to characterizing optimal strategies for performing
mental rotation. We found that an adaptive, nonparametric model
performs best at the classic rotation task \cite{Shepard1971}, as it
is able to actively monitor the confidence of its estimate and
intelligently choose the direction of rotation. Two simpler models
based on heuristics performed much worse: they did not maintain the
linear relationship with the true angle of rotation due to lack of
robustness against local maxima or incorrect rotation direction. The
nonparametric model qualitatively approximates human behavioral
judgments, both in linearity \cite{Shepard1971} and variability
\cite{Just1976}. Future work will collect empirical data from
participants to perform a more thorough, quantitative analysis.  We
conclude that from this initial survey, models of mental rotation
which take an ``active'', directed approach seem well-suited to
explaining human behavior in such metacognitive domains as mental
simulation.

% One option for improving the performance of the simple models would be
% to fit a threshold value, below which local maxima would be
% ignored. However, this strategy would be rather brittle, for if the
% distribution of shapes changed, the threshold would have to be
% re-learned. It is possible that people exhibit this behavior, but we
% cannot make any assumptions one way or the other without empirical
% data.

% A further strength of the nonparametric model over the other models,
% however, is that it is likely to generalize well to three
% dimensions. One aspect of \cite{Osborne:2012tm} which we have not yet
% explored is their main contribution of \textit{active sampling}, in
% which samples are iteratively selected to maximally decrease the
% expected variance of $Z$. Due to the sequential constraint of mental
% rotation, this strategy is not particularly useful in two dimensions,
% as there are only every two directions in which to rotate. In three
% dimensions, however, there are an infinite number of directions that
% could be chosen after every step.
